Level-Set method on \texttt{p4est}

### 2014-08-29 paper/tex/introduction.tex
More update on the introduction

\item \textbf{What is the main challenge for parallelization?}
One of the main challenges in parallelizing level-set algorithms on adaptive grids, is handling the grid itself in parallel. One option is to replicate the entire grid on each processor and use off-the-self graph partitioners, e.g. Metis or ParMetis \fcite{metis/parmetis paper}, for load balancing and domain decomposition \fcite{origial deal.II paper}. This approach, however, is only scalable to a few hundred processor at best and is limited by the size of the grid itself. Moreover, use of a general-purpose graph partitioner adds extra overhead that limits the overall scalability even further. Interestingly, Quadtrees and Octrees have nice spatial ordering that naturally lead to the concept of Space Filling Curves (SFC) \fcite{SFC reference} which can be efficiently exploited for high quality load balancing.

\item \textbf{Parallel octrees: p4est, dendro, octor, etc. How do we plan to use p4est?}
The idea of using SFCs for parallel partitioning of Quad-/Octrees is not new in itself. In \fcite{octor} authors first used the idea of using Morton curve (also known as Z-curve) for traversing the leaves of an Octree which is then used to index and distribute them between processors. Dendro \fcite{dendro paper} follows similar ideas for developing a parallel geometric multigrid on balanced octrees. More recently, authors in \fcite{p4est} extended these ideas to a collection, or a ``forest'', of octrees that are connected through a common, potentially unstructured, hexahedral grid. They then use a global Morton curve to partition the entire forest in parallel. Implementation of these algorithms, which were shown to scale to more than 200,000 processors, are publicly available through simple API provided by the \texttt{p4est} library \fcite{p4est website?}. In fact the algorithms presented in this paper are directly implemented on top of \texttt{p4est} library and we do not discuss any algorithm that is already covered by \texttt{p4est} implementation. Instead make appropriate references to \fcite{p4st paper} whenever necessary.

@@

\item \textbf{Parallel semi-lagrangian methods? What are the challenges? Whats out there?}

>>>>

\item \textbf{What is the main challenge for parallelization?}
One of the main challenges in parallelizing level-set algorithms on adaptive grids, is handling the grid itself in parallel. One option is to replicate the entire grid on each processor and use off-the-self graph partitioners, e.g. Metis or ParMetis \fcite{metis/parmetis paper} or Zoltan \fcite{zoltan paper}, for load balancing and domain decomposition \fcite{origial deal.II paper}. This approach, however, is only scalable to a few hundred processor at best and is limited by the size of the grid itself. Moreover, use of a general-purpose graph partitioner adds extra overhead that limits the overall scalability even further. Interestingly, Quadtrees and Octrees have nice spatial ordering that naturally lead to the concept of Space Filling Curves (SFC) \fcite{SFC reference} which can be efficiently exploited for high quality load balancing.

\item \textbf{Parallel octrees: p4est, dendro, octor, etc. How do we plan to use p4est?}
The idea of using SFCs for parallel partitioning of Quad-/Octrees is not new in itself and has been tried by many researchers. For instance, Octor \fcite{octor} uses a Morton curve (also known as Z-curve) for traversing the leaves of an Octree which is then used to index and distribute them between processors and has been shown to scale up to 6200 cores \fcite{mantle convection octor}. Dendro \fcite{original dendro paper} is another example of parallel octree code in which similar ideas are used for parallel partitioning and development of a parallel geometric multigrid that has been scaled up to about 32000 cores \fcite{dendro parallel multigrid}. More recently, authors in \fcite{p4est} extended these ideas to a collection, or a ``forest'', of octrees that are connected through a common, potentially unstructured, hexahedral grid. They then use a global Morton curve to partition the entire forest in parallel. Implementation of these algorithms, which were shown to scale to more than 200,000 processors, are publicly available through simple API provided by the \texttt{p4est} library \fcite{p4est website?}. In fact the algorithms presented in this paper are directly implemented on top of \texttt{p4est} library and we do not discuss any algorithm that is already covered in \texttt{p4est paper}. Instead we focus on parallelization of level-set method on Quad-/Octrees and make appropriate references to \fcite{p4st paper} whenever necessary.

\item \textbf{Parallel semi-lagrangian methods? What are the challenges? Whats out there?}
Parallel level-set algorithms can be categorized in three separate groups: 1) Parallel advection algorithms, 2) Parallel reinitialization algorithms, and 3) Parallel motion under curvature algorithms. In this article we focus on developing algorithms for the first two categories. We also make use of the popular \texttt{PETSc} \fcite{petsc} library for solution of the linear systems that arise from the third category. Moreover, \texttt{PETSc} provides primitives, such as parallel ghosted vector and scatter/gather operations, which simplifies the implementation.

Eulerian advection schemes can easily be parallelized but unfortunately are limited by the CFL condition. For adaptive grids, this could be very restrictive as the tree could be locally refined very deeply to capture quantities of interest. Semi-Lagrangian methods combine the unconditional stability of Lagrangian methods and the ease of use of Eulerian grids and have been successfully used for advecting the level-set function on Quad-/Octree grids before \fcite{Frederic's paper}. However, parallelizing the Semi-Lagrangian algorithm in a domain decomposition context is not an easy task. The reason for this is twofold: First, depending on the CFL number, the departure points may end up outside the ghost region and in remote processors that are potentially far away. This requires a very dynamic and nonuniform communication pattern which is complicated to implement. For adaptive grids situation is even more complicated due to the asymmetric nature of communication, i.e. every processor knows the message destination but no processor knows which processor to expect message from and how many. Second, depending on the CFL number, the distribution of departure points could be entirely independent of the original parallel partitioning. As a result large CFL numbers could result in very load-unbalanced distribution of departure points which restricts the scalability of the algorithm. Both of these problems, of course, could be avoided by choosing $\text{CFL} \le 1$ but that would defeat the purpose of using Semi-Lagrangian algorithm in the first place.

@@
